Implications of Measurement Errors in Water Use Estimates for Water Management

### 2024-08-12 Codes/1_analysis.qmd
manuscript writing started

## Prices and costs
```{r}
corn_price <- 6 # (bu/acre)
e_price <- 0.15 # ($/kwh)
w_price <- e_price * 500 / 12 # ($/inch)
other_cost <- 800 # ($/acre)
```

@@

```{r}
#| eval: false
system_data <-
  par_set %>%
  rowwise() %>%
  dplyr::mutate(farmer_data = list(
    gen_farmer_data(par_set, num_farmers)
  )) %>%
  dplyr::mutate(prod_data = list(
    #--- generate production data ---#
    gen_prod_data(farmer_data)
  ))

@@

@@

# Results Presentation

@@

```{r, eval = FALSE}
#--- proxy ---#
results_p <- readRDS("Results/results_proxy.rds")

@@

farm_data <-
  data.table(system_data)[prod_het == "medium", farmer_data][[1]] %>%
  #--- use farmer 1 ---#
  .[farmer == 1, ] %>%
  .[, farmer := NULL] %>%
  expand_grid_df(., data.table(w = seq(0, 15, by = 0.1))) %>%
  .[, yield := ymax * (1 - alpha * exp(-beta * w))] %>%
  .[, w_cost := w_price * w] %>%
  .[, profit := corn_price * yield - w_cost - other_cost] %>%
  .[, marginal_cost_w := w_price] %>%
  .[, marginal_prod_w := ymax * alpha * beta * exp(-beta * w)] %>%
  .[, marginal_profit_w := marginal_prod_w * corn_price - marginal_cost_w]

opt_w <- unique(farm_data$opt_w)

quota_data <-
  data.table(
    farmer = c(1, 2, 1, 2),
    measure_type = c("Water", "Water", "Proxy", "Proxy"),
    quota = c(7, 7, 4, 10)
  ) %>%
  .[, w := quota] %>%
  .[quota > opt_w, w := opt_w] %>%
  farm_data[., on = .(w), roll = "nearest"]

pi_dif_2 <- quota_data[farmer == 2 & measure_type == "Proxy", profit] - quota_data[farmer == 2 & measure_type == "Water", profit]
pi_dif_1 <- quota_data[farmer == 1 & measure_type == "Proxy", profit] - quota_data[farmer == 1 & measure_type == "Water", profit]
```

The red dotted lines indicate the effective water use quota. Under the water quota system, both farmers receive 7 inches of water quota. Since the unconstrained profit maximizing water use is `r round(opt_w, digits = 2)` inches, both farmers use up all the allocated quota and use 10 inches of water. Under the proxy quota system, they are allocated the same amount of proxy quota. However, measurement error in proxy quota resulted in effective water quota of 4 and 10 inches for farmer A and B, respectively. Both farmers use their allocated amout of water. Under both proxy and water quota systems, the total water use is 14 inches. Farmer B makes `r pi_dif_2` less under the water quota system compared to the proxy quota system. However, Farmer A makes `r -pi_dif_1` more under the water quota system compared to the proxy quota system. This comes from an important nature of crop production with respsect to water: declining marginal impact of water on crop yield. As can be seen in the profit-water relationship, removing an inch of water from 6 inches has much larger negative impact compared to revmoing an inch of water from 10 inches.

```{r}
#| warning: false
(
  g_illustration_homogeneous <-
    ggplot() +
    geom_line(
      data = farm_data,
      aes(y = profit, x = w)
    ) +
    geom_vline(
      data = quota_data,
      aes(xintercept = quota),
      linetype = 2,
      color = "red"
    ) +
    geom_point(
      data = quota_data,
      aes(y = profit, x = w),
      color = "blue",
      size = 2
    ) +
    facet_grid(measure_type ~ farmer) +
    ylim(0, NA) +
    xlab("Water Use (inches)") +
    ylab("Profit ($/acre)") +
    theme_bw()
)

@@

farm_data <-
  data.table(system_data)[prod_het == "medium", farmer_data][[1]] %>%
  .[, dev_from_7 := abs(opt_w - 7)] %>%
  .[, dev_from_13 := abs(opt_w - 13)] %>%
  .[(dev_from_7 == min(dev_from_7) | dev_from_13 == min(dev_from_13)), ] %>%
  expand_grid_df(., data.table(w = seq(0, 15, by = 0.1))) %>%
  .[, yield := ymax * (1 - alpha * exp(-beta * w))] %>%
  .[, w_cost := w_price * w] %>%
  .[, profit := corn_price * yield - w_cost - other_cost] %>%
  .[, marginal_cost_w := w_price] %>%
  .[, marginal_prod_w := ymax * alpha * beta * exp(-beta * w)] %>%
  .[, marginal_profit_w := marginal_prod_w * corn_price - marginal_cost_w] %>%
  .[, farmer_txt := "A"] %>%
  .[opt_w == max(opt_w), farmer_txt := "B"]

@@

quota_data_het <-
  data.table(
    farmer_txt = c("A", "B", "A", "B"),
    measure_type = c("Water", "Water", "Proxy", "Proxy"),
    quota = c(7, 7, 4, 10)
  ) %>%
  opt_w[., on = .(farmer_txt)] %>%
  .[, w := quota] %>%
  .[quota > opt_w, w := opt_w] %>%
  setnames("farmer_txt", "farmer_id") %>%
  nest_by(farmer_id) %>%
  dplyr::mutate(merged = list(
    farm_data[farmer_txt == farmer_id, ][data.table(data), on = .(w), roll = "nearest"]  
  )) %>%
  dplyr::select(farmer_id, merged) %>%
  unnest(cols = c(merged)) %>%
  data.table() %>%
  .[, .(farmer, farmer_txt, w, opt_w, profit, measure_type, quota)]

@@

```{r}
(
  g_profit_quota_comp_stat <-
    ggplot(data = quota_plot_data_comp_stat) +
    geom_line(aes(y = profit, x = w, color = measure_type)) +
    facet_grid(u_dev ~ prod_het) +
    ylab("Profit ($/acre)") +
    xlab("Realized Water Use (inches)")
)
```

@@

@@

### Trading reduces welfare loss associated with the use of proxy

@@

(
  g_trade_effect_profit <-
    ggplot(data = trading_plot_data) +
    geom_line(aes(y = profit, x = w, color = type)) +
    facet_grid(u_dev ~ prod_het) +
    scale_color_brewer(palette = "Set1") +
    ylab("Profit ($/acre)") +
    xlab("Realized Water Use (inches)") +
    scale_color_discrete(name = "Measurement Type") +
    theme_bw() +
    theme(legend.position = "bottom")
)

ggsave("Writing/figures/g_trade_effect_profit.pdf", g_trade_effect_profit, height = 3.5, width = 6)
```

### Trading reduces uncertainty in water use slack

@@

## Profit Distribution

@@

#! Nearest matching based on water use (not target water use).

@@

profit_ratio_quota <- 
  readRDS(here::here("Results/profit_ratio_quota.rds")) %>%
  .[dir_cor == "0" & ratio > 0 & ratio < 3, ] %>%
  .[, dev_w := abs(w-8), by = .(prod_het, u_dev)] %>%
  .[, .SD[dev_w == min(dev_w)], by = .(prod_het, u_dev)] %>%
  .[, .(u_dev, prod_het, ratio, prob)] %>%
  .[, instrument_type := "Quota"]

@@

ggsave("Writing/figures/g_profit_ratio.pdf", g_profit_ratio, height = 3.5, width = 6)
```

>>>>

## Prices and costs

```{r}
corn_price <- 6 # (bu/acre)
e_price <- 0.15 # ($/kwh)
w_price <- e_price * 500 / 12 # ($/inch)
other_cost <- 800 # ($/acre)

save(corn_price, e_price, w_price, other_cost)
```

@@

set.seed(812614)

system_data <-
  par_set %>%
  rowwise() %>%
  dplyr::mutate(farmer_data = list(
    gen_farmer_data(par_set, num_farmers)
  )) %>%
  dplyr::mutate(prod_data = list(
    #--- generate production data ---#
    gen_prod_data(farmer_data)
  ))

@@

```{r}
g_opt_w_dist <-
  system_data %>%
  dplyr::select(prod_het, farmer_data) %>%
  unnest(cols = c(farmer_data)) %>%
  data.table() %>%
  .[, type_txt := case_when(
    prod_het == "zero" ~ "Heterogeneity: Zero",
    prod_het == "medium" ~ "Heterogeneity: Medium",
    prod_het == "high" ~ "Heterogeneity: High"
  )] %>%
  .[, type_txt := factor(
    type_txt,
    levels = c("Heterogeneity: Zero", "Heterogeneity: Medium", "Heterogeneity: High")
  )] %>%
  .[opt_w_true < 30, ] %>%
  ggplot(data = .) +
  geom_histogram(
    aes(
      x = opt_w_true,
      y = ifelse(after_stat(count) > 0, after_stat(count), NA)
    ),
    color = "blue",
    fill = NA
  ) +
  facet_grid(type_txt ~ .) +
  xlab("Optimal Water Use (inches)") +
  ylab("Count") +
  scale_x_continuous(breaks = seq(0, 30, by = 2)) +
  theme_bw()

ggsave("Writing/figures/g_opt_w_dist.pdf", g_opt_w_dist, height = 5, width = 6)
```

@@

# Results

@@

```{r, eval = FALSE}
#| eval: false
#--- proxy ---#
results_p <- readRDS("Results/results_proxy.rds")

@@

w_A <- 7
w_B <- 7
p_A <- 4
p_B <- 10

farm_data_1 <-
  data.table(system_data)[prod_het == "medium", farmer_data][[1]] %>%
  #--- use farmer 1 ---#
  .[farmer == 1, ] %>%
  .[, farmer := NULL] %>%
  expand_grid_df(., data.table(w = seq(0, 15, by = 0.1))) %>%
  .[, yield := ymax * (1 - alpha * exp(-beta * w))] %>%
  .[, w_cost := w_price * w] %>%
  .[, profit := corn_price * yield - w_cost - other_cost] %>%
  .[, marginal_cost_w := w_price] %>%
  .[, marginal_prod_w := ymax * alpha * beta * exp(-beta * w)] %>%
  .[, marginal_profit_w := marginal_prod_w * corn_price - marginal_cost_w]

opt_w <- unique(farm_data_1$opt_w)

quota_data <-
  data.table(
    farmer = c("A", "B", "A", "B"),
    measure_type = c("Water", "Water", "Proxy", "Proxy"),
    quota = c(w_A, w_B, p_A, p_B)
  ) %>%
  .[, w := quota] %>%
  .[quota > opt_w, w := opt_w] %>%
  farm_data_1[., on = .(w), roll = "nearest"]

pi_dif_B <- quota_data[farmer == "B" & measure_type == "Proxy", profit] - quota_data[farmer == "B" & measure_type == "Water", profit]
pi_dif_A <- quota_data[farmer == "A" & measure_type == "Proxy", profit] - quota_data[farmer == "A" & measure_type == "Water", profit]
```

The red dotted lines indicate the effective water use quota. Under the water quota system, both farmers receive 7 inches of water quota. Since the unconstrained profit maximizing water use is `r round(opt_w, digits = 2)` inches, both farmers use up all the allocated quota and use 10 inches of water. Under the proxy quota system, they are allocated the same amount of proxy quota. However, measurement error in proxy quota resulted in effective water quota of 4 and 10 inches for farmer A and B, respectively. Both farmers use their allocated amout of water. Under both proxy and water quota systems, the total water use is 14 inches. Farmer B makes `r pi_dif_B` less under the water quota system compared to the proxy quota system. However, Farmer A makes `r -pi_dif_A` more under the water quota system compared to the proxy quota system. This comes from an important nature of crop production with respsect to water: declining marginal impact of water on crop yield. As can be seen in the profit-water relationship, removing an inch of water from 6 inches has much larger negative impact compared to revmoing an inch of water from 10 inches.

```{r}
#| warning: false
(
  g_illustration_homogeneous <-
    ggplot() +
    geom_line(
      data = farm_data_1,
      aes(y = profit, x = w)
    ) +
    geom_vline(
      data = quota_data,
      aes(xintercept = quota),
      linetype = 2,
      color = "red"
    ) +
    geom_point(
      data = quota_data,
      aes(y = profit, x = w),
      color = "blue",
      size = 2
    ) +
    facet_grid(measure_type ~ farmer) +
    ylim(0, NA) +
    xlab("Water Use (inches)") +
    ylab("Profit ($/acre)") +
    theme_bw()
)

@@

w_A <- 7
w_B <- 7
p_A <- 4
p_B <- 10

farm_data <-
  data.table(system_data)[prod_het == "medium", farmer_data][[1]] %>%
  .[, dev_from_7 := abs(opt_w - 7)] %>%
  .[, dev_from_13 := abs(opt_w - 13)] %>%
  .[(dev_from_7 == min(dev_from_7) | dev_from_13 == min(dev_from_13)), ] %>%
  expand_grid_df(., data.table(w = seq(0, 15, by = 0.1))) %>%
  .[, yield := ymax * (1 - alpha * exp(-beta * w))] %>%
  .[, w_cost := w_price * w] %>%
  .[, profit := corn_price * yield - w_cost - other_cost] %>%
  .[, marginal_cost_w := w_price] %>%
  .[, marginal_prod_w := ymax * alpha * beta * exp(-beta * w)] %>%
  .[, marginal_profit_w := marginal_prod_w * corn_price - marginal_cost_w] %>%
  .[, farmer_txt := "A"] %>%
  .[opt_w == max(opt_w), farmer_txt := "B"] %>%
  .[, marginal_profit_p := ifelse(farmer_txt == "A", marginal_profit_w / w_A * p_A, marginal_profit_w / w_B * p_B)]

@@

quota_data_het <-
  data.table(
    farmer_txt = c("A", "B", "A", "B"),
    measure_type = c("Water", "Water", "Proxy", "Proxy"),
    quota = c(w_A, w_B, p_A, p_B)
  ) %>%
  opt_w[., on = .(farmer_txt)] %>%
  .[, w := quota] %>%
  .[quota > opt_w, w := opt_w] %>%
  setnames("farmer_txt", "farmer_id") %>%
  nest_by(farmer_id) %>%
  dplyr::mutate(merged = list(
    farm_data[farmer_txt == farmer_id, ][data.table(data), on = .(w), roll = "nearest"]
  )) %>%
  dplyr::select(farmer_id, merged) %>%
  unnest(cols = c(merged)) %>%
  data.table() %>%
  .[, .(farmer, farmer_txt, w, opt_w, profit, measure_type, quota)]

@@

```{r}
(
  g_profit_quota_comp_stat <-
    ggplot(data = quota_plot_data_comp_stat) +
    geom_line(aes(y = profit, x = w, color = measure_type)) +
    facet_grid(u_dev ~ prod_het) +
    ylab("Profit ($/acre)") +
    xlab("Realized Water Use (inches)") +
    scale_color_discrete(name = "Measurement Type") +
    theme_bw() +
    theme(legend.position = "bottom")
)

ggsave("Writing/figures/g_profit_quota_comp_stat.pdf", g_profit_quota_comp_stat, height = 4.5, width = 6)
```

@@

# Quota Trading

@@

## Trading reduces welfare loss associated with the use of proxy

@@

(
  g_trade_effect_profit <-
    ggplot(data = trading_plot_data) +
    geom_line(aes(y = profit, x = w, color = type)) +
    facet_grid(u_dev ~ prod_het) +
    scale_color_brewer(palette = "Set1") +
    ylab("Profit ($/acre)") +
    xlab("Realized Water Use (inches)") +
    scale_color_discrete(name = "Measurement Type") +
    guides(color = guide_legend(nrow = 2, byrow = TRUE)) +
    theme_bw() +
    theme(legend.position = "bottom")
)

ggsave("Writing/figures/g_trade_effect_profit.pdf", g_trade_effect_profit, height = 4.5, width = 6)
```

## Trading reduces uncertainty in water use slack

@@

## Trading illustration

Illustrate the effect of allowing proxy trading. Water trading does not happen as it is always optimal.

```{r}
farm_data_A <-
  copy(farm_data_1) %>%
  .[, farmer := "A"] %>%
  .[, marginal_profit_p := marginal_profit_w / w_A * p_A]

farm_data_B <-
  copy(farm_data_1) %>%
  .[, farmer := "B"] %>%
  .[, marginal_profit_p := marginal_profit_w / w_B * p_B]

mpp_w_A <- gam(w ~ s(marginal_profit_p), data = farm_data_A)
mpp_w_B <- gam(w ~ s(marginal_profit_p), data = farm_data_B)

w_data <-
  data.table(
    marginal_profit_p = seq(0, min(farm_data_A[, max(marginal_profit_p)], farm_data_B[, max(marginal_profit_p)]), length = 1001)
  ) %>%
  .[, w_A := predict(mpp_w_A, newdata = .)] %>%
  .[, w_B := predict(mpp_w_B, newdata = .)] %>%
  .[, tot_w := w_A + w_B] %>%
  .[, dev := abs(tot_w - 14)] %>%
  .[dev == min(dev), ]

pi_w_A <- gam(profit ~ s(w), data = farm_data_A)
pi_w_B <- gam(profit ~ s(w), data = farm_data_B)

proxy_trade_data_A <-
  data.table(w = w_data$w_A) %>%
  .[, profit := predict(pi_w_A, newdata = .)] %>%
  .[, farmer := "A"]

proxy_trade_data_B <-
  data.table(w = w_data$w_B) %>%
  .[, profit := predict(pi_w_B, newdata = .)] %>%
  .[, farmer := "B"]

proxy_trade_data <-
  rbind(proxy_trade_data_A, proxy_trade_data_B) %>%
  .[, measure_type := "Proxy"]

g_illustration_homogeneous +
  geom_point(data = proxy_trade_data, aes(x = w, y = profit)) 
```

# Profit Distribution

@@

# ! Nearest matching based on water use (not target water use).

@@

profit_ratio_quota <-
  readRDS(here::here("Results/profit_ratio_quota.rds")) %>%
  .[dir_cor == "0" & ratio > 0 & ratio < 3, ] %>%
  .[, dev_w := abs(w - 8), by = .(prod_het, u_dev)] %>%
  .[, .SD[dev_w == min(dev_w)], by = .(prod_het, u_dev)] %>%
  .[, .(u_dev, prod_het, ratio, prob)] %>%
  .[, instrument_type := "Quota"]

@@

ggsave("Writing/figures/g_profit_ratio.pdf", g_profit_ratio, height = 5, width = 6)
```

### 2024-08-12 Writing/manuscript.Rmd
manuscript writing started

```{r profit-quota-water-het, fig.cap = "Comparison of profit from agricultural production between the water and quota systems"}
(
  g_profit_quota_w_vs_p <-
    sum_res %>%
    .[instrument_type == "Quota", ] %>%
    .[(dir_cor == "0" & u_dev == 0.3 & measure_type == "Proxy") | measure_type == "Water", ] %>%
    ggplot(data = .) +
    geom_line(aes(y = profit, x = w, color = measure_type)) +
    facet_grid(. ~ prod_het) +
    ylab("Target Water Use (inches)") +
    xlab("Realized Water Use (inches)") +
    scale_color_discrete(name = "Measurement Type") +
    theme_bw() +
    theme(
      legend.position = "bottom"
    )
)
```

@@

```{r increase-me-degree, fig.cap = "The impacts of an increase in the degree of measurement error on profit"}
q_udev <- sum_res[(dir_cor == "0" & measure_type == "Proxy"), ]
w_udev <-
  sum_res[measure_type == "Water", ] %>%
  .[, u_dev := NULL] %>%
  expand_grid_df(., data.table(u_dev = c(0.1, 0.3, 0.5)))

quota_plot_data_comp_stat <-
  rbind(q_udev, w_udev) %>%
  .[instrument_type == "Quota", ]

(
  g_profit_quota_comp_stat <-
    ggplot(data = quota_plot_data_comp_stat) +
    geom_line(aes(y = profit, x = w, color = measure_type)) +
    facet_grid(u_dev ~ prod_het) +
    ylab("Profit ($/acre)") +
    xlab("Realized Water Use (inches)") +
    scale_color_discrete(name = "Measurement Type") +
    theme_bw() +
    theme(legend.position = "bottom")
)
```

@@

Figure below presents the ratio of profit under quota-based relative to water-based system for quota (red) and permit trading (blue). Ratio of less than 1 indicates making less profit under the quota-based system compared to the water-based system.

@@

@@

```{r, fig.height = 5}
system_data %>%
  dplyr::select(prod_het, farmer_data) %>%
  unnest(cols = c(farmer_data)) %>%
  data.table() %>%
  .[, type_txt := case_when(
    prod_het == "zero" ~ "Heterogeneity: Zero",
    prod_het == "medium" ~ "Heterogeneity: Medium",
    prod_het == "high" ~ "Heterogeneity: High"
  )] %>%
  .[, type_txt := factor(
    type_txt,
    levels = c("Heterogeneity: Zero", "Heterogeneity: Medium", "Heterogeneity: High")
  )] %>%
  .[opt_w_true < 30, ] %>%
  ggplot(data = .) +
  geom_histogram(
    aes(
      x = opt_w_true,
      y = ifelse(after_stat(count) > 0, after_stat(count), NA)
    ),
    color = "blue",
    fill = NA
  ) +
  facet_grid(type_txt ~ .) +
  xlab("Optimal Water Use (inches)") +
  ylab("Count") +
  scale_x_continuous(breaks = seq(0, 30, by = 2)) +
  theme_bw()
```

>>>>

```{r profit-quota-water-het, fig.cap = "Comparison of profit from agricultural production between the water and quota systems"}
knitr::include_graphics("figures/g_profit_quota_w_vs_p.pdf")
```

@@

```{r increase-me-degree, fig.cap = "The impacts of an increase in the degree of measurement error on profit"}

knitr::include_graphics("figures/g_profit_quota_comp_stat.pdf")
```

@@

Figure \@ref(fig:distributional-equity) presents the cumulative probability of the ratio of profit under quota-based relative to water-based system for quota (red) and permit trading (blue). Ratio of less than 1 indicates making less profit under the quota-based system compared to the water-based system. If farmers make just as much profir under proxy-based regulation compared to water-based regulation, then the cumulative probability line would be a vertical line at 1. For example, under the high degree of measurement error and medium heterogenenity scenario, some farmers would make about the half of what they would make under the water quota system. At the lower end of the distribution are farmers who would be constrained severely by water use quota, but whose water use estiamtes are much higher than their actual water use. At the higher end of the distribution are farmers who would be constrained by water use quota, but whose water use estiamtes are much lower than their actual water use. Famres who suffer significantly under the quota system can develop strong dissatisfaction of the quota-based regulation system, which in turn may lead to installing water meter themselves. If they discover the over-estimation of their water use, they may appeal to legal actions against the administrator for their unfair treatment. Allowind quota trading greatly diminishes the distributional inequality, though not completely. This is because unfortunate farmers who suffered from over-estimation of their water use can now at least purchase proxy quota to increase their production level.

@@

## Limitations

Assumptions.
Temporal

@@

```{r dist-opt-w, fig.cap = "Distribution of optimal water use under no water use constraints", fig.height = 5}
knitr::include_graphics("figures/g_opt_w_dist.pdf")
```
